Neural models for computer-assisted musical orchestration: a preliminary study

### 2020-05-13 paper/latex/paper.tex
Add paper.bib

\subsection{Baseline}
In order to have a baseline to compare our results against, we attempted to solve the classification problem using various parametric classifiers. The classifiers we tested were Support Vector Machine (SVM), Random Forest, and K-Nearest Neighbors. We used the implementations provided in the scikit-learn library for each classifier. For SVM, we used SVC with an RBF kernel. For Random Forest, we set the maximum depth of each tree to be 15. Each classifier was wrapped in a MultiOutputClassifier to achieve the multi-label nature of this problem. We found SVM to have the highest accuracy of the three classifiers across all experiments. All of the following baselin experiments used 50,000 generated samples with a train-test split of 60/40. Each sample is a combination of one or more instruments and is four seconds in length. The features used are the mel frequency cepstral coefficients (MFCCs) of the resulting combination, with a total of 100 coefficients per sample.

@@

The first deep model we tested was a Convolutional Neural Network (CNN). CNN shows good performance on classification assignment for its strong ability to extract spacial features, so we implemented a basic CNN with three convolutional layers and two fully connected layers. Each convolutional layer is followed by a BatchNorm layer, a ReLU activation layer and a 2 * 2  MaxPool layer with a stride of 2. The kernel size is 3 * 3 with a stride of 1 and a padding of 1. The number of filters are 8, 16, 32. This setting is simple but we can make a good comparison between it and the baseline. \\

@@

Given that audio is one kind of sequential data, it is easy to consider Long Short-Term Memory network(LSTM) as test model. In our setting, we added one LSTM layer with 32 outputs units following three convolutional layers following the same pattern as the CNN described earlier. After LSTM layer, we reshaped the output into the same size as the output of the last convulational layer, which was fed to another convolutional layer with the same structure and kernel size, yielding a tensor of dimensions $32\times 8 \times 21$. Finally, we flattened out the outputs and fed them into fully a connected layer with Dropout layer, followed by another fully connected layer. Finally, the sigmoid function was applied to the final layer in order to get a probability between 0 and 1 for each individual label. Since there is no competition relationship between different classes for each class can be classification result equally, it is reasonable for us to take sigmoid activations and used binary classification for each class to treat them independently. We minimize the binary cross entropy(BCE) loss between the predictions and targets. We implemented our neural networks with PyTorch and trained them with Adam optimizer. The model architecture is shown in Figure\ref{cnn_lstm}.\\

@@

% For bibtex users:
\bibliography{ISMIRtemplate}

>>>>

\subsection{Baseline}
In order to have a baseline to compare our results against, we attempted to solve the classification problem using various parametric classifiers. The classifiers we tested were Support Vector Machine (SVM), Random Forest, and K-Nearest Neighbors. We used the implementations provided in the scikit-learn library for Python. For SVM, we used SVC with an RBF kernel. For Random Forest, we set the maximum depth of each tree to be 15. Each classifier was wrapped in a MultiOutputClassifier to achieve the multi-label nature of this problem. We found SVM to have the highest accuracy of the three classifiers across all experiments. All of the following baseline experiments used 50,000 generated samples with a train-test split of 60/40. Each sample is a combination of one or more instruments and is four seconds in length. The features used are the mel frequency cepstral coefficients (MFCCs) of the resulting combination, with a total of 100 coefficients per sample. We chose to use MFCCs for this setting because using the mel spectrogram leads to a number of features that is too high for the parametric classifiers.

@@

The first deep model we tested was a Convolutional Neural Network (CNN). CNNs show good performance on classification problems for their ability to extract spatial features. \textbf{Reference needed}. Our architecture contained three convolutional layers and two fully connected layers. Each convolutional layer is followed by a BatchNorm layer, a ReLU activation layer and a $2 \times 2$  MaxPool layer with a stride of 2. The kernel size is $3 \times 3$ with a stride of 1 and a padding of 1. The number of filters are 8, 16, 32.

@@

Given the sequential nature of audio, a Long Short-Term Memory (LSTM) network provides a way to learn long term dependencies in the data \cite{LSTM}.

 In our setting, we added one LSTM layer with 32 outputs units following three convolutional layers following the same pattern as the CNN described earlier. After LSTM layer, we reshaped the output into the same size as the output of the last convulational layer, which was fed to another convolutional layer with the same structure and kernel size, yielding a tensor of dimensions $32\times 8 \times 21$. Finally, we flattened out the outputs and fed them into fully a connected layer with Dropout layer, followed by another fully connected layer. Finally, the sigmoid function was applied to the final layer in order to get a probability between 0 and 1 for each individual label. Since there is no competition relationship between different classes for each class can be classification result equally, it is reasonable for us to take sigmoid activations and used binary classification for each class to treat them independently. We minimize the binary cross entropy(BCE) loss between the predictions and targets. We implemented our neural networks with PyTorch and trained them with Adam optimizer. The model architecture is shown in Figure\ref{cnn_lstm}.\\

@@

% For bibtex users:
\bibliography{paper}
